Vagueness, Noise, and Signaling

### 2015-08-12 paper/version_04/paper.tex
explain new dynamics
dynamics write up

Signaling agents can adapt their dispositions to act, given some feedback about their past
success, in multiple ways. Usually, we would assume that changes in behavior should, at least
on average, lean towards increasing chances of communicative success. Such a population of
signaling agents and their behavioral adaptations can be described at different levels of
abstraction. At the level of individual agents, we can picture a more or less idealized process
of how each agent adapts dispositions for future actions based on various pieces of information
available to the agent, e.g., about the game, the other players' behavior and the like. More
abstractly, at the level of a population of agents, we can describe how average behavioral
dispositions will evolve. The population-level perspective abstracts over petty stochastic
fluctuations and zooms in on the general tendency or direction of evolution that ensues from
some (often: simplified) behavior at the agent-level.

To better understand the interaction of selective pressure successful communication and
stochastic noise, we look at a population-level dynamic that describes the most likely
evolutionary path of a population of signaling agents who imitate other agents' behavior but
may occasionally confuse states for one another. We call our dynamic \emph{imprecise
  conditional imitation dynamic} (\icid), because it derives from an update protocol based on
conditional imitation and allows for imprecise perception and realization of states. In the
special limiting case where imprecision vanishes, the process is just the well-known replicator
dynamic \citep{TaylorJonker1978:Evolutionary-St}. To appreciate the \icid and its motivation,
we first briefly review the replicator dynamic and its derivation from conditional imitation of
successes.

@@

The \emph{expected utility} of choices at each choice point is:
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
    \States} \frac{\Pr(\state) \cdot \Sstrat(\messg \probbar \state)}{\sum_{\state'} \Pr(\state') \cdot \Sstrat(\messg \probbar \state')} \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
outcomes, weighted by how likely they are to occur. For instance, the expected utility for the
sender of choosing message $\messg \in \Messgs$, given tate $\state \in \States$ and a receiver
strategy $\Rstrat$ is the sum, for each act $\act \in \Acts$, of the probability that the
receiver will choose act $\act$ given message $\messg$ times the actual utility of using act
$\act$ for state $\state$, given $\state$ and $\messg$.

@@

For numerical simulations, it is often more convenient to work with a discrete-time
formulation, which maps current strategies $\Sstrat$ and $\Rstrat$ to strategies $\RD(\Sstrat)$
and $\RD(\Rstrat)$ that describe the population at the next discrete time step. Like the
continuous-time version, the discrete-time replicator dynamics tracks changes in frequency of
choices as proportional to their expected utilities
\citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
  utilities are always positive.}
\begin{align}
  \label{eq:RDDiscrete}
  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
    \EU(\messg, \state,\Rstrat)} {\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot
   \EU(\messg', \state,\Rstrat)} \\
    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
    \EU(\act, \messg,\Sstrat)} {\sum_{\act'} \Rstrat(\act' \probbar \messg)
    \cdot \EU(\act', \messg,\Sstrat)}  \,. \nonumber
\end{align}
Intuitively put, this definition updates the average frequency with which each choice (message
or act) will occur, independently at each choice point (state or message). For a given choice
point, say a state $\state$, the probability of seeing $\messg$ played by an average agent in
the population \emph{before} the update is $\Sstrat(\messg \probbar \state)$. \emph{After} the
update it will be proportional to the frequency before the update times the expected utility of
$\messg$ at state $\state$. The denominator is then a normalizing constant to make sure that we
obtain probabilities that sum to one. In particular, the denominator in the
definition above captures the average expected utility of all choices at the relevant choice
point.

@@

The replicator dynamic is an abstract population-level dynamic that describes the mean expected
change of behavioral dispositions in a population of signalers. There are several ways of
deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
here on one of the simplest: \emph{conditional imitation of success} \citep[see][for this and
many other derivations]{Sandholm2010:Population-Game}. We choose this agent-level derivation,
because it is more convenient than its alternatives when we want to add imprecision in the
following. The main idea is that agents imitate the behavior of another agent at some choice
point with a probability that is proportional to the expected utility of the latter agent's act
for that choice point. Since the sim-max games that we will be looking at in
Section~\ref{sec:exploring-rdd} have positive utilities, we may even identify the switching
probability with the expected utility directly.

Let's consider sender strategies, as the receiver case is parallel (as long as there is no
noise or imprecision). Revision by imitation of success proceeds as follows. Call
(misleadingly!) the agent who gets a chance to change behavior ``learner'' and the possibly
to-be-imitated agent ``teacher.'' A random learner is drawn from the population and given a
chance to change behavior at choice point $\state$. The probability that our learner plays
$\messg$ is $\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled
teacher does at $\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar
\state)$. The learner then starts using $\messg'$ instead of $\messg$ with probability
$\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and $\messg$ could be the same.)

If agents get repeated updating chances for their choice points, the \emph{expected change} of
frequency for $\messg$ at $\state$ becomes:
\begin{align}
  \label{eq:MeanChange}
  \dot{\Sstrat}(\messg \probbar \state) = P(\messg' \rightarrow m, \state) - P(\messg \rightarrow m', \state)\,,
\end{align}
where $P(\messg' \rightarrow m, \state)$ is the ``inflow'' probability that agents switch from
$\messg'$ to $\messg$ and $P(\messg \rightarrow m', \state)$ is the ``outflow'' probability
that agents switch from $\messg$ to $\messg'$. Since we are dealing with expectations in a huge
population, these probabilities can be given as:
\begin{align*}
  P(\messg' \rightarrow m, \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
    \state)}_{\text{learner plays $\messg'$}} \cdot
  \underbrace{\Sstrat(\messg \probbar \state)}_{\text{teacher plays $\messg$}} \cdot
  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} \\
  P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
    \state)}_{\text{learner plays $\messg$}} \cdot
  \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
  \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
\end{align*}

From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
to:
\begin{align*}
  \dot{\Sstrat}(\messg \probbar \state) & = \Sstrat(\messg \probbar \state) \cdot \EU(\messg,
  \state, \Rstrat) - \Sstrat(\messg \probbar \state) \cdot  \sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
& = \Sstrat(\messg \probbar \state) \cdot \Big ( \EU(\messg, \state,\Rstrat) - \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
  \state,\Rstrat) \Big )\,,
\end{align*}
which is the continuous-time formulation of the replicator dynamic in (\ref{eq:RDContinuous}).

@@

The above derivation of the replicator dynamic assumes that agents can discriminate choices and
choice points perfectly. Let's dispense with that assumption. With an eye towards sim-max
games, we will assume that receivers choose interpretation actions $\Acts = \States$ and that
any two states may have a chance of being confused for one another, proportional to how similar
they are. This will affect the way that agents realize the (pure) strategies that they carry
and also how learners observe teacher choices and the probability with which they switch to the
observed teacher choice.

Agents carry pure dispositions to act, but noise can affect their realization. As a sender,
every agent maps states to messages: these are subjectively perceived states, and no longer
necessarily also the actually occurring state. As a receiver, every agent maps messages to
state interpretations: these are intended interpretations that need not always be faithfully
realized. This means that behavioral strategies $\Sstrat \in (\Delta(\Messgs))^\States$ and
$\Rstrat \in (\Delta(\States))^\Messgs$ represent the average proportions of \emph{actual}
behavioral dispositions in the population, the realization and observation of which can be
distorted by confusion of similar states.

If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
probability that a given agent observes state $\mystate{o}$. States $\mystate{a}$ and
$\mystate{o}$ may be identical. In the limiting case, $P_o(\mystate{o} \probbar \mystate{a})$
could be one only if $\mystate{a}$ and $\mystate{o}$ are identical. For the more general case,
we should treat $P_o(\mystate{o} \probbar \mystate{a})$ as a primitive of the model. Similarly,
if a given receiver intends to select interpretation $\mystate{i}$, we will assume that the
probability with which state $\mystate{r}$ is realized is
$P_r(\mystate{r} \probbar \mystate{i})$. Section~\ref{sec:exploring-rdd} gives a concrete
specification of $P_o$ and $P_r$ in terms of similarity of states.

A single round of play is then governed by five pieces of stochastic information, where
previously there were only three. At the outset, the actual state is chosen with probability
$P_a(\mystate{a})$. A randomly sampled sender observes $\mystate{o}$ with probability
$P_o(\mystate{o} \probbar \mystate{a})$ and so we expect $\messg$ with probability
$\Sstrat(\messg \probbar \mystate{o})$. (Senders respond to observations, not the actual
state!) As perception of messages is flawless, we expect a randomly sampled receiver to intend
to realize interpretation $\mystate{i}$ with probability
$\Rstrat(\mystate{i} \probbar \messg)$.  The actually realized interpretation will be
$\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$.

Expected utilities of choices at choice points should take into account as well that actual
states need not be observed states, and intended interpretations need not be realized
interpretations. First, note that 
\begin{align*}
  P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) = \frac{P_a(\mystate{a}) P_o(\mystate{o}
    \probbar \mystate{a})}{\sum_{\state}P_a(\state) P_o(\mystate{o}
    \probbar \state)}
\end{align*}
is the probability that $\mystate{a}$ is actual if $\mystate{o}$ is observed. The probability
that a random sender produces $\messg$ when the actual state is $\mystate{a}$ is:
\begin{align*}
  P_{\Sstrat}(\messg \probbar \mystate{a}) = \sum_{\mystate{o}} P_o(\mystate{o} \probbar
  \mystate{a}) \Sstrat(\messg \probbar \mystate{o})\,. 
\end{align*}
The probability that the actual state is $\mystate{a}$ if a random sender produced $\messg$ is:
\begin{align*}
  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) = \frac{P_a(\mystate{a})
  P_{\Sstrat}(\messg \probbar \mystate{a})}{\sum_{\state} P_a(\state)
  P_{\Sstrat}(\messg \probbar \state)}\,. 
\end{align*}
The probability that $\mystate{r}$ is realized by a random receiver in response to
message $\messg$ is:
\begin{align*}
  P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
  \mystate{i}) \Rstrat(\mystate{i} \probbar \messg)\,.
\end{align*}
This lets us capture the expected utilities for observed states (sender) and intended
interpretations (receiver) by taking into consideration what the likely actual states and
realized interpretations will be:
\begin{align*}
  \EU(\messg , \mystate{o}, \Rstrat) & = \sum_{\mystate{a}}  P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \sum_{\mystate{r}}  P_\Rstrat(\mystate{r} \probbar
  \messg) \utils(\mystate{a}, \messg, \mystate{r})\,, \\
  \EU(\mystate{i} , \messg, \Sstrat) & = \sum_{\mystate{a}}
  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \sum_{\mystate{r}}
  P_{r}(\mystate{r} \probbar \mystate{i})  \utils(\mystate{a}, \messg, \mystate{r})\,.
\end{align*}
If the conditional probabilities $P_o$ and $P_r$ are trivial, i.e., assign probability 0 to the
observability and realizability of non-identical states, above definitions reduce to the
previous definitions of expected utilities. This also legitimates the overload of notation.

@@

As before, suppose that senders receive a chance to change their behavior independently for a
given choice point. In the present case, this would be a chance to change how to respond to a
subjectively perceived state $\mystate{o}$, which need not be the actual one. We must therefore
take into account what the probability is that, given that the learner observed $\mystate{o}$,
he will observe a randomly sampled teacher play some $\messg$ or other. The probability
$P_o(\messg \probbar \mystate{o})$ that some teacher plays $\messg$ in the same situation in
which some learner observes $\mystate{o}$ is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as
defined above):
\begin{align*}
  P_o(\messg \probbar \mystate{o}) = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \myts P_{\Sstrat}(\messg \probbar \mystate{a})\,.
\end{align*}
The probability $P(\messg' \rightarrow \messg, \state)$ that a randomly sampled learner switches
from $\messg'$ to $\messg$ in subjectively perceived state $\mystate{o}$ is therefore:
\begin{align*}
  P(\messg' \rightarrow \messg, \mystate{o}) = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
    \mystate{o})}_{\text{learner plays $\messg'$ at $\mystate{o}$}} \cdot
  \underbrace{P_{o}(\messg \probbar \mystate{o})}_{\text{observe teacher play $\messg$}} \cdot
  \underbrace{\EU(\messg, \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
\end{align*}
The mean change to the proportion of $\messg$ choices at state $\state$ are then as before:
\begin{align*}
  \dot{\Sstrat}(\messg \probbar \mystate{o}) & = P(\messg' \rightarrow m, \mystate{o}) - P(\messg
  \rightarrow m', \mystate{o}) \\
  & = \sum_{\messg'} \Sstrat(\messg' \probbar \mystate{o}) \myts P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \mystate{o}) \myts
  P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat) \\
  & = P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \Sstrat(\messg \probbar \mystate{o}) \myts \sum_{\messg'} P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat)
\end{align*}
This has an elegant and practical discrete-time solution:\footnote{To see how the discrete-time formulation
  gives rise to the continuous-time formulation, let's assume that update steps are
  infinitesimally small, so that
  \begin{align*}
    \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
    \probbar \state)\\ 
    &= \frac{P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat)
      - \Sstrat(\messg \probbar \state) \sum_{\messg'} P_o(\messg' \probbar \state) \myts
      \EU(\messg', \state, \Rstrat) }{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts
      \EU(\messg', \state, \Rstrat)}\,.
  \end{align*}
By dropping the nominator, which is constant for all $\messg$ for fixed
$\state$, we obtain the above continuous-time formulation.}
\begin{align*}
  \Sstrat'(\messg \probbar \state) = \frac{P_o(\messg \probbar \state) \myts \EU(\messg,
  \state, \Rstrat)}{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts \EU(\messg',
  \state, \Rstrat)}\,.
\end{align*}

>>>>

Signaling agents can adapt their dispositions to act, given some feedback about their past
success, in multiple ways. Usually, we would assume that changes in behavior should, at least
on average, lean towards increasing chances of communicative success. Such behavioral
adaptations can be described at different levels of abstraction. At the level of individual
agents, we can picture a more or less idealized process of how each agent adapts dispositions
for future actions based on various pieces of information available to the agent, e.g., about
the game, the other players' behavior and the like. More abstractly, at the level of a
population of agents, we can describe how average behavioral dispositions will evolve. The
population-level perspective abstracts over petty stochastic fluctuations and zooms in on the
general tendency or direction of evolution that ensues from some (often: simplified) behavior
at the agent-level.

To better understand the interaction of confusability of stimuli and selective pressure towards
successful communication, we look at a population-level dynamic that describes the most likely
evolutionary path of a population of signaling agents who imitate other agents' behavior but
may occasionally confuse states for one another. We call our dynamic \emph{imprecise
  conditional imitation dynamic} (\icid), because it derives from an update protocol based on
conditional imitation and allows for imprecise perception and realization of states. In the
special limiting case where imprecision vanishes, the process is just the well-known replicator
dynamic \citep{TaylorJonker1978:Evolutionary-St}. To appreciate the \icid and its motivation,
we first briefly review the replicator dynamic and its derivation from conditional imitation of
successes.

@@

The \emph{expected utility} of choices at each choice point is:
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
    \States} \frac{\Pr(\state) \cdot \Sstrat(\messg \probbar \state)}{\sum_{\state'} \Pr(\state') \cdot \Sstrat(\messg \probbar \state')} \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
outcomes, weighted by how likely they are to occur. For instance, the expected utility for the
sender of choosing message $\messg \in \Messgs$, given tate $\state \in \States$ and a receiver
strategy $\Rstrat$ is the sum, for each act $\act \in \Acts$, of the probability that the
receiver will choose act $\act$ given message $\messg$ times the actual utility of using act
$\act$, given $\state$ and $\messg$.

@@

For numerical simulations, it can be more convenient to work with a discrete-time formulation,
which maps current strategies $\Sstrat$ and $\Rstrat$ to strategies $\Sstrat'$ and $\Rstrat'$
that describe the population at the next discrete time step. Like the continuous-time version,
the discrete-time replicator dynamics tracks changes in frequency of choices as proportional to
their expected utilities \citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The
  formulation given in Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one
  we will be looking at in Section~\ref{sec:exploring-rdd}, where utilities are always
  non-negative and expected utilities are always positive.}\textsuperscript{,}\footnote{The
  notation $\propto$, for ``proportional to'', means that the right-hand side must possibly yet
  be normalized so as to have probabilities sum to one. In general, writing $P(x) \propto f(x)$
  for any function $f \mycolon X \rightarrow \mathds{R}$ is shorthand for
  $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. The implicit normalizing denominators for the
  replicator equations are average expected utilities, e.g., for the sender:
  $\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg', \state,\Rstrat)$.}
\begin{align}
  \label{eq:RDDiscrete}
  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \cdot
    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\act \probbar \messg) & \propto \Rstrat(\act \probbar \messg) \cdot
    \EU(\act, \messg,\Sstrat)\,.   
\end{align}
For a given choice point, say a state $\state$, the probability of seeing $\messg$ played by an
average agent in the population \emph{after} the update is proportional to the probability of
seeing it \emph{before} the update, which is $\Sstrat(\messg \probbar \state)$, times the
expected utility of $\messg$ at state $\state$. Intuitively put, frequencies of choices change
by a gradient of current frequency and a measure of how good they are.

@@

The replicator dynamic is an abstract population-level dynamic that describes the mean expected
change of behavioral dispositions in a population of signalers. There are several ways of
deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
here on one of the simplest: \emph{conditional imitation of success} \citep[see][for this and
many other derivations]{Sandholm2010:Population-Game}. The main idea is that agents imitate the
behavior of another agent at some choice point with a probability that is proportional to the
expected utility of the latter agent's act for that choice point. Since the sim-max games that
we will be looking at in Section~\ref{sec:exploring-rdd} have positive utilities, we may
identify the switching probability with the expected utility.

Let's consider sender strategies, as the receiver case is parallel. Revision by imitation of
success proceeds as follows. Call (misleadingly!) the agent who gets a chance to change
behavior ``learner'' and the possibly to-be-imitated agent ``teacher.'' A random learner is
drawn from the population and given a chance to change behavior at choice point $\state$. The
probability that our learner plays $\messg$ is $\Sstrat(\messg \probbar \state)$. The learner
observes what a randomly sampled teacher does at $\state$. That would be $\messg'$ with
probability $\Sstrat(\messg' \probbar \state)$. The learner then starts using $\messg'$ instead
of $\messg$ with probability $\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and
$\messg$ could be the same.)

If agents get repeated update chances for their choice points, the \emph{expected change} of
frequency for $\messg$ at $\state$ becomes:
\begin{align}
  \label{eq:MeanChange}
  \dot{\Sstrat}(\messg \probbar \state) = P(\messg' \rightarrow m, \state) - P(\messg \rightarrow m', \state)\,,
\end{align}
where $P(\messg' \rightarrow m, \state)$ is the ``inflow'' probability that agents switch from
$\messg'$ to $\messg$ and $P(\messg \rightarrow m', \state)$ is the ``outflow'' probability
that agents switch from $\messg$ to $\messg'$. Since we are dealing with expectations in a huge
population, the probability $P(\messg' \rightarrow m, \state)$ of switching from $\messg'$ to
$\messg$ at $\state$ can be written as:
\begin{align*}
  P(\messg' \rightarrow m, \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
    \state)}_{\text{learner plays $\messg'$}} \cdot
  \underbrace{\Sstrat(\messg \probbar \state)}_{\text{teacher plays $\messg$}} \cdot
  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} 
  % P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
  %   \state)}_{\text{learner plays $\messg$}} \cdot
  % \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
  % \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
\end{align*}
From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
to:
\begin{align*}
  \dot{\Sstrat}(\messg \probbar \state) & = \Sstrat(\messg \probbar \state) \cdot \EU(\messg,
  \state, \Rstrat) - \Sstrat(\messg \probbar \state) \cdot  \sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
& = \Sstrat(\messg \probbar \state) \cdot \Big ( \EU(\messg, \state,\Rstrat) - \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
  \state,\Rstrat) \Big )\,,
\end{align*}
which is the continuous-time formulation of the replicator dynamic in (\ref{eq:RDContinuous}).

@@

The above derivation of the replicator dynamic assumes that agents can discriminate choices and
choice points perfectly. Let's dispense with that assumption. With an eye towards sim-max
games, we will assume that receivers choose interpretation actions $\Acts = \States$ and that
any two states, but not messages, may be confused for one another.\footnote{It is relatively
  straightforward to also incorporate confusability of messages, but this is irrelevant to our
  present purposes.} Confusability of states will affect how agents behave (on average), how
they perceive others that they might want to imitate, and how good any behavioral disposition
actually is.

To keep matters simple, let us assume that agents carry pure dispositions to act, but noise can
affect their realization. As a sender, every agent maps states to messages: these are
subjectively perceived states, and no longer necessarily also the actually occurring state. As
a receiver, every agent maps messages to state interpretations: these are intended
interpretations that need not always be faithfully realized. This means that behavioral
strategies $\Sstrat \in (\Delta(\Messgs))^\States$ and $\Rstrat \in (\Delta(\States))^\Messgs$
represent the average proportions of \emph{actual} behavioral dispositions in the population,
the realization and observation of which can be distorted by confusion of similar states.

If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
probability that a given agent observes state $\mystate{o}$. Similarly, if a given receiver
intends to select interpretation $\mystate{i}$, we will assume that the probability with which
state $\mystate{r}$ is realized is $P_r(\mystate{r} \probbar
\mystate{i})$.
Section~\ref{sec:exploring-rdd} gives a concrete specification of $P_o$ and $P_r$ in terms of
similarity of states. For the time being, it is best to treat $P_o$ and $P_r$ as arbitrary
conditional confusion probabilities.

A single round of play of a sim-max game is then governed by five pieces of stochastic
information, where previously there were only three. At the outset, the actual state is chosen
with probability $P_a(\mystate{a})$. A randomly sampled sender observes $\mystate{o}$ with
probability $P_o(\mystate{o} \probbar \mystate{a})$ and so we expect $\messg$ with probability
$\Sstrat(\messg \probbar \mystate{o})$. (Senders respond to observations, not the actual
state!) As perception of messages is flawless, we expect a randomly sampled receiver to intend
to realize interpretation $\mystate{i}$ with probability
$\Rstrat(\mystate{i} \probbar \messg)$.  The actually realized interpretation will be
$\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$.

Expected utilities of choices at choice points should also take into account that actual states
need not be observed states, and intended interpretations need not be realized
interpretations. First, note that
\begin{align*}
  P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) \propto P_a(\mystate{a}) P_o(\mystate{o}
    \probbar \mystate{a})
\end{align*}
is the probability that $\mystate{a}$ is actual if $\mystate{o}$ is observed. The probability
that a random sender produces $\messg$ when the actual state is $\mystate{a}$ is:
\begin{align*}
  P_{\Sstrat}(\messg \probbar \mystate{a}) = \sum_{\mystate{o}} P_o(\mystate{o} \probbar
  \mystate{a}) \Sstrat(\messg \probbar \mystate{o})\,. 
\end{align*}
The probability that the actual state is $\mystate{a}$ if a random sender produced $\messg$ is:
\begin{align*}
  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \propto P_a(\mystate{a})
  P_{\Sstrat}(\messg \probbar \mystate{a})\,. 
\end{align*}
The probability that $\mystate{r}$ is realized by a random receiver in response to
message $\messg$ is:
\begin{align*}
  P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
  \mystate{i}) \Rstrat(\mystate{i} \probbar \messg)\,.
\end{align*}
This lets us capture the expected utilities for observed states (sender) and intended
interpretations (receiver) by taking into consideration what the likely actual states and
realized interpretations will be:
\begin{align*}
  \EU(\messg , \mystate{o}, \Rstrat) & = \sum_{\mystate{a}}  P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \sum_{\mystate{r}}  P_\Rstrat(\mystate{r} \probbar
  \messg) \utils(\mystate{a}, \messg, \mystate{r})\,, \\
  \EU(\mystate{i} , \messg, \Sstrat) & = \sum_{\mystate{a}}
  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \sum_{\mystate{r}}
  P_{r}(\mystate{r} \probbar \mystate{i})  \utils(\mystate{a}, \messg, \mystate{r})\,.
\end{align*}
If the conditional probabilities $P_o$ and $P_r$ are trivial, i.e., assign probability 0 to the
observability and realizability of non-identical states, above definitions reduce to the
previous definitions of expected utilities. This also legitimates the overload of notation.

@@

As before, suppose that senders receive a chance to change their behavior independently for a
given choice point. In the present case, this would be a chance to change how to respond to a
subjectively perceived state $\mystate{o}$, which need not be the actual one. We must then
consult the probability $P_o(\messg \probbar \mystate{o})$ that, given that the learner
observed $\mystate{o}$, he will simultaneously observe a randomly sampled teacher play
$\messg$. This is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as defined above):
\begin{align*}
  P_o(\messg \probbar \mystate{o}) = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \myts P_{\Sstrat}(\messg \probbar \mystate{a})\,.
\end{align*}
The probability $P(\messg' \rightarrow \messg, \state)$ that a randomly sampled learner switches
from $\messg'$ to $\messg$ in subjectively perceived state $\mystate{o}$ is therefore:
\begin{align*}
  P(\messg' \rightarrow \messg, \mystate{o}) = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
    \mystate{o})}_{\text{learner plays $\messg'$ at $\mystate{o}$}} \cdot
  \underbrace{P_{o}(\messg \probbar \mystate{o})}_{\text{observe teacher play $\messg$}} \cdot
  \underbrace{\EU(\messg, \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
\end{align*}
The mean change to the proportion of $\messg$ choices at state $\state$ are then:
\begin{align}
  \label{eq:nRDSender}
  \dot{\Sstrat}(\messg \probbar \mystate{o}) & = P(\messg' \rightarrow m, \mystate{o}) - P(\messg
  \rightarrow m', \mystate{o}) \nonumber \\
  & = \sum_{\messg'} \Sstrat(\messg' \probbar \mystate{o}) \myts P_o(\messg \probbar
  \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \mystate{o}) \myts
  P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat) \nonumber \\
  & = P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \Sstrat(\messg \probbar \mystate{o}) \myts \sum_{\messg'} P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat)\,.
\end{align}

The case of the receiver is mostly analogous. Presented with an update opportunity for choice
point $\messg$, a learner will observe a random teacher choose interpretation $\mystate{o}$
with probability:\footnote{We assume that the noise $P_o(\mystate{o} \probbar \mystate{a})$
  that applies to the sender's observations also applies to the receiver's observations when
  trying to imitate other agents' strategies. We could easily make the model more complex and
  introduce another measure of noise for state confusability during receiver attempts to
  imitate. We refrain from it here, because we see no immediate theoretical gain.}
\begin{align*}
  P_o(\mystate{o} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
  \myts P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
  \myts \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
  \mystate{i}) \myts \Rstrat(\mystate{i} \probbar \messg) \,.
\end{align*}
The probability $P(\state' \rightarrow \state, \messg)$ that a randomly sampled learner switches
from using $\state'$ to $\state$ in response to $\messg$ is therefore:
\begin{align*}
  P(\state' \rightarrow \state, \messg) = \sum_{\state'} \underbrace{\Rstrat(\state' \probbar
    \messg)}_{\text{learner plays $\state'$}} \cdot
  \underbrace{P_{o}(\state \probbar \messg)}_{\text{observe teacher play $\state$}} \cdot
  \underbrace{\EU(\state, \messg, \Rstrat)}_{\text{EU teacher choice}}
\end{align*}
Parallel to those for the sender case, this gives rise to:
\begin{align}
  \label{eq:nRDReceiver}
  \dot{\Rstrat}(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg)
  \myts P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'}
  \Rstrat(\state \probbar \messg) \myts
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \nonumber \\
  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar
  \messg) \myts \sum_{\state'}
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) 
\end{align}
